---
title: "Graphing a Handful of Vanguard Funds"
author: "Rick Dean"
format: 
  html:
    toc: false
    toc-depth: 4
    toc-location: "right"
    number-sections: true
    number-offset: 0
    self-contained: true
    smooth-scroll: true
    code-fold: true
    code-block-bg: "#f1f3f5"
    code-block-border-left: "#31BAE9"
    code-overflow: wrap
    tbl-cap-location: "bottom"
    fig-width: 12
    fig-height: 6
    fig-align: "center"
    fig-cap-location: "bottom"
    css: ../style.css
    link-external-newwindow: true
    abstract-title: "Abstract"
    abstract: "Taking a handful of Vanguard mutual index funds, perform data wrangling and graphing of adjusted closing prices using the R language."
---

```{r}
#| warning: false
#| message: false

library(data.table)
library(ggplot2)
library(quantmod)
library(magrittr)
library(purrr)
library(RplotterPkg)
library(RstocksPkg)
```

# Graphing a Handful of Vanguard Funds

## Vanguard Index Fund Closing Prices

:::topic
Select the Vanguard index funds.
:::

The following funds are selected for graphing.

* Bonds:
    + [VBILX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vbilx) -- Inter-Term Bond Index Fund Admiral Shares
    + [VBIRX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vbirx) -- Vanguard Short-Term Bond Index Fund Admiral Shares

* Growth:
    + [VIMAX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vimax) -- Vanguard Mid-Cap Index Fund Admiral Shares
    + [VFIAX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vfiax) -- Vanguard 500 Index Fund Admiral Shares
    + [VIGAX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vigax) -- Vanguard Growth Index Fund Admiral Shares
    + [VTSAX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vtsax) -- Vanguard Total Stock Market Index Fund Admiral Shares

* Balanced-Income-Value:
    + [VWELX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vwelx) -- Vanguard Wellington Fund Investor Shares
    + [VWINX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vwinx) -- Vanguard Wellesley Income Fund Investor Shares
    + [VVIAX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vviax) -- Vanguard Value Index Fund Admiral Shares

* Sector-Foreign:
    + [VGSLX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vgslx) -- Vanguard Real Estate Index Fund Admiral Shares
    + [VGHCX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vghcx) -- Vanguard Health Care Fund Investor Shares
    + [VFWAX](https://investor.vanguard.com/investment-products/mutual-funds/profile/vfwax) -- Vanguard FTSE All-World ex-US Index Fund Admiral Shares

:::topic
Show the fund symbols and define start/end times of interest.
:::

Create named vectors of fund symbols and start/end times:

```{r}
symbols_v <- c(
  VBILX = "Inter-term Bond Index Adm-R",
  VIMAX = "Mid-Cap Index Fund Adm-R",
  VFIAX = "500 Index Fund Adm-R",
  VIGAX = "Growth Index Fund Adm-R",
  VTSAX = "Total Stock Market Index Adm-R",
  VVIAX = "Value Index Fund Adm-R",
  VGSLX = "Real Estate Index Fund-RN",
  VWELX = "Wellington Fund Inv",
  VFWAX = "FTSE All World ex-US Adm",
  VGHCX = "Health Care Fund Inv",
  VBIRX = "Short-Term Bond Index Adm-N",
  VWINX = "Wellesley Income Fund Inv-RN"
)

times_v <- c(
  start = "2019-01-01",
  end = "2022-12-31"
)
```

:::topic
Get the historic closing prices for all the funds.
:::

```{r}
#| warning: false

fund_prices_dt <- RstocksPkg::get_stock_prices(
  symbols = names(symbols_v),
  from = times_v[["start"]],
  to = times_v[["end"]],
  periodicity = "weekly",
  warnings = FALSE
)
```

## Growth Plots

:::topic
Define the `data.table` for growth funds.
:::

```{r}
growth_symbols_v <- c(
  VIMAX = "Mid-Cap Index Fund Adm-R",
  VFIAX = "500 Index Fund Adm-R",
  VIGAX = "Growth Index Fund Adm-R",
  VTSAX = "Total Stock Market Index Adm-R"
)
growth_dt <- fund_prices_dt[Symbol %in% names(growth_symbols_v),]
```

```{r}
#| fig-height: 7
#| fig-width: 12
#| fig-cap: Standardized adjusted closing prices for Vanguard Growth Index Funds

build_plot <- function(id, dt, symbols, show_x, show_y){
  symbol <- names(symbols)[[id]]
  plot_dt <- dt[Symbol == symbol, ]
  a_mean <- mean(plot_dt$Adjusted)
  a_sd <- sd(plot_dt$Adjusted)
  plot_dt[, Z := (Adjusted-a_mean)/a_sd]
  
  hide_x_tics <- TRUE
  hide_y_tics <- TRUE
  if(id %in% show_x){
    hide_x_tics <- FALSE
  }
  if(id %in% show_y){
    hide_y_tics <- FALSE
  }
  
  RplotterPkg::create_scatter_plot(
    df = plot_dt,
    aes_x = "Date",
    aes_y = "Z",
    subtitle = symbols[[id]],
    y_limits = c(-3, 3),
    y_major_breaks = seq(-3, 3, 1.0),
    hide_x_tics = hide_x_tics,
    hide_y_tics = hide_y_tics,
    connect = TRUE,
    show_minor_grids = FALSE,
    show_pts = FALSE
  )
}

growth_plots_lst <- purrr::map(
  1:4, 
  build_plot, 
  dt = growth_dt, 
  symbols = growth_symbols_v,
  show_x = c(3, 4),
  show_y = c(1, 3)
)

layout <- list(
  plots = growth_plots_lst,
  rows = c(1, 1, 2, 2),
  cols = c(1, 2, 1, 2)
)

RplotterPkg::multi_panel_grid(
  layout = layout,
  row_heights = c(8, 8.5),
  col_widths = c(10.5,10)
)
```

## Balanced-Income-Value Plots

```{r}
balanced_symbols_v <- c(
  VWELX = "Wellington Fund Inv",
  VWINX = "Wellesley Income Fund Inv-RN",
  VVIAX = "Value Index Fund Adm-R"
)
balanced_dt <- fund_prices_dt[Symbol %in% names(balanced_symbols_v),]
```

```{r}
#| fig-height: 7
#| fig-width: 12
#| fig-cap: Standardized adjusted closing prices for Vanguard Balanced/Income/Value Index Funds

balanced_plots_lst <- purrr::map(
  1:3, 
  build_plot, 
  dt = balanced_dt, 
  symbols = balanced_symbols_v,
  show_x = c(2, 3),
  show_y = c(1, 3)
)

layout <- list(
  plots = balanced_plots_lst,
  rows = c(1, 1, 2),
  cols = c(1, 2, 1)
)

RplotterPkg::multi_panel_grid(
  layout = layout,
  row_heights = c(8, 8.5),
  col_widths = c(10.5,10)
)
```

## Bonds Plots

```{r}
bond_symbols_v <- c(
  VBILX = "Inter-term Bond Index Adm-R",
  VBIRX = "Short-Term Bond Index Adm-N"
)
bond_dt <- fund_prices_dt[Symbol %in% names(bond_symbols_v),]
```


```{r}
#| fig-height: 6
#| fig-width: 12
#| fig-cap: Standardized adjusted closing prices for Vanguard Bond Index Funds

bond_plots_lst <- purrr::map(
  1:2, 
  build_plot, 
  dt = bond_dt, 
  symbols = bond_symbols_v,
  show_x = c(1, 2),
  show_y = 1
)

layout <- list(
  plots = bond_plots_lst,
  rows = c(1, 1),
  cols = c(1, 2)
)

RplotterPkg::multi_panel_grid(
  layout = layout,
  row_heights = c(8, 8.5),
  col_widths = c(10.5,10)
)
```

## Sector-Foreign Plots

```{r}
sector_symbols_v <- c(
  VGSLX = "Real Estate Index Fund-RN",
  VGHCX = "Health Care Fund Inv",
  VFWAX = "FTSE All World ex-US Adm"
)
sector_dt <- fund_prices_dt[Symbol %in% names(sector_symbols_v),]
```

```{r}
#| fig-height: 7
#| fig-width: 12
#| fig-cap: Standardized adjusted closing prices for Vanguard Sector/Foreign Index Funds

sector_plots_lst <- purrr::map(
  1:3, 
  build_plot, 
  dt = sector_dt, 
  symbols = sector_symbols_v,
  show_x = c(2, 3),
  show_y = c(1, 3)
)

layout <- list(
  plots = sector_plots_lst,
  rows = c(1, 1, 2),
  cols = c(1, 2, 1)
)

RplotterPkg::multi_panel_grid(
  layout = layout,
  row_heights = c(8, 8.5),
  col_widths = c(10.5,10)
)
```
