---
title: "Tech Dividends, Part 1"
author: "Rick Dean"
format: 
  html:
    toc: false
    toc-depth: 4
    toc-location: "right"
    number-sections: true
    number-offset: 0
    self-contained: true
    smooth-scroll: true
    code-fold: true
    code-block-bg: "#f1f3f5"
    code-block-border-left: "#31BAE9"
    code-overflow: wrap
    tbl-cap-location: "bottom"
    fig-width: 12
    fig-height: 6
    fig-align: "center"
    fig-cap-location: "bottom"
    css: ../style.css
    link-external-newwindow: true
    abstract-title: "Abstract"
    abstract: "The following notes/R scripts were inspired from an article [Tech Dividends, Part 1](https://rviews.rstudio.com/2019/08/07/tech-dividends-part1/) by Jonathan Regenstein."
---

```{r}
#| warning: false
#| message: false

library(data.table)
library(ggplot2)
library(here)
library(magrittr)
library(quantmod)
library(RplotterPkg)
library(RstocksPkg)
```

::: note
Note: Be aware that this application has an internet request for dividend data that will take time to complete.
:::

# Tech Dividends, Part 1

## The data.

:::topic
NASDAQ company symbols
:::

Read the symbols of NASDAQ companies with market cap greater than $10B.  The source is
[Stock Screener](https://www.nasdaq.com/market-activity/stocks/screener). We only want companies 
that have IPO'ed greater than 2007.

```{r}
symbols_file_path <- file.path(here::here(), "data", "nasdaq_mega_large_medium_small.csv")
nasdaq_symbols_dt <- data.table::fread(symbols_file_path) %>% 
  .[, .(Symbol, Name, `IPO Year`, `Market Cap`, Sector)] %>% 
  .[`IPO Year` > 2007,] %>% 
  .[order(-`Market Cap`)]
```


:::topic
Dividend prices for companies over the last 10 years.
:::

We want to get dividend history for the top 100 NASDAQ companies based on "Market Cap".
```{r}
#| warning: false

top_nasdaq_symbols <- nasdaq_symbols_dt$Symbol[1:100]
nasdaq_dividends_dt <- RstocksPkg::get_dividends(
  symbols = top_nasdaq_symbols,
  from = "2013-01-01",
  to = "2022-12-31"
) %>% 
  na.omit()
```

Add a *year* variable to our dividends data.table.
```{r}
#| fig-cap: Format of current data.table with 491 NASDAQ observations posting dividends

nasdaq_dividends_dt[, Year := as.numeric(format(Date, "%Y"))]
RplotterPkg::create_table(
  x = nasdaq_dividends_dt[1:6,],
  container_width_px = 400
)
```

## First date when the companies posted a dividend.

Get the first observation from each *Symbol* grouping where a dividend was posted.

```{r}
#| tbl-cap: First Posted Dividend Dates for Top NASDAQ Companies(21 x 4)

first_dividend_dt <- nasdaq_dividends_dt[, .SD[1L], by = Symbol]

RplotterPkg::create_table(
  x = first_dividend_dt[1:6,],
  container_width_px = 400
)
```

We have a total of 21 companies with their first dividend posting dates.

## Count for each year when a company first granted a dividend.

```{r}
#| tbl-cap: Yearly number of companies that granted a dividend (8 x 2)

first_dividends_year_count_dt <- first_dividend_dt[, .N, by = Year][order(Year)]

RplotterPkg::create_table(
  x = first_dividends_year_count_dt,
  container_width_px = 200
)
```

## Most recent posting in each company.

Get the most recent dividend from each company:

```{r}
#| tbl-cap: Most recent posting of dividends in each company(21 x 4)

latest_dividends_dt <- nasdaq_dividends_dt[, .SD[.N], by = Symbol][order(-Dividend)]

RplotterPkg::create_table(
  x = latest_dividends_dt,
  container_width_px = 400
)
```

Plot the most recent dividends posted in 2022.

```{r}
#| warning: false
#| fig-cap: Last Dividends Paid by NASDAQ Companies in year 2022

RplotterPkg::create_scatter_plot(
  df = latest_dividends_dt,
  aes_x = "Date",
  aes_y = "Dividend",
  aes_label = "Symbol",
  aes_label_size = 3,
  x_title = "Week (2022)",
  y_title = "Dividends($)",
  rot_y_tic_label = TRUE,
  x_limits = c(as.Date("2022-11-01"), as.Date("2022-12-31")),
  x_major_date_breaks = "1 week",
  x_date_labels = "%b%e",
  y_limits = c(0.0, 5.0),
  y_major_breaks = seq(0.0, 5.0, 1.0),
  pts_size = 1.2,
  pts_color = "blue",
  pts_fill = "blue",
  show_minor_grids = F
)
```

## Plot all the dividends posted by each company

The above data.table *nasdaq_dividends_dt* has 491 observations from the top 100 NASDAQ companies from 2013-01-01 to 2022-12-31. Create a scatter plot of all these companies. 

```{r}
#| fig-cap: All Dividends Paid by NASDAQ Companies from 2013-01-01 to 2022-12-31

RplotterPkg::create_scatter_plot(
  df = nasdaq_dividends_dt,
  aes_x = "Date",
  aes_y = "Dividend",
  aes_fill = "Symbol",
  x_title = "Year",
  y_title = "Dividends($)",
  rot_y_tic_label = TRUE,
  x_limits = c(as.Date("2013-01-01"), as.Date("2022-12-31")),
  x_major_date_breaks = "1 year",
  x_date_labels = "%Y",
  y_limits = c(0.0, 5.5),
  y_major_breaks = seq(0.0,5.5,0.5),
  pts_size = 1.2,
  show_minor_grids = F,
  show_legend = F
) 
```

## Grouped statistical summaries

:::topic
Dividend stats by company
:::

Compute descriptive statistics on *Dividend* for the 21 top companies from *nasdaq_dividends_dt*. 

```{r}
#| tbl-cap: Dividend Statistics for top NASDAQ Companies (21 x 7)

stats_comp_dividends_dt <- nasdaq_dividends_dt[, .(N = .N, Mean = mean(Dividend), Median = median(Dividend),  Min = min(Dividend), Max = max(Dividend), Total = sum(Dividend)), by = Symbol][order(-Total)]

RplotterPkg::create_table(
  x = stats_comp_dividends_dt,
  container_width_px = 600
)
```

This is an interesting table where for example we can see that the ticker "AGNC" posted dividends 107 times with an average amount of 0.21 and a total of 22.55 over the 2013 to 2022 time frame.

:::topic
Dividend stats by year
:::

```{r}
#| tbl-cap:  Dividend Statistics across the years for top NASDAQ Companies (7 x 7)

stats_year_dividends_dt <- nasdaq_dividends_dt[, .(N = .N, Mean = mean(Dividend), Median = median(Dividend), Min = min(Dividend), Max = max(Dividend), Total = sum(Dividend)), by = Year][order(Year)]

RplotterPkg::create_table(
  x = stats_year_dividends_dt,
  container_width_px = 600
)
```

Evidently 2022 was a good year for dividends where if we held the 21 companies, then we would receive a total of 84 postings for a total of $49.18.
