---
title: "Tech Dividends, Part 2"
author: "Rick Dean"
format: 
  html:
    toc: false
    toc-depth: 4
    toc-location: "right"
    number-sections: true
    number-offset: 0
    self-contained: true
    smooth-scroll: true
    code-fold: true
    code-block-bg: "#f1f3f5"
    code-block-border-left: "#31BAE9"
    code-overflow: wrap
    tbl-cap-location: "bottom"
    fig-width: 12
    fig-height: 6
    fig-align: "center"
    fig-cap-location: "bottom"
    css: ../style.css
    link-external-newwindow: true
    abstract-title: "Abstract"
    abstract: "The following notes/R scripts were inspired from an article [Tech Dividends, Part 2](https://rviews.rstudio.com/2019/08/17/tech-dividends-part-2/) by Jonathan Regenstein."
---

# Tech Dividends, Part 2

```{r}
#| warning: false
#| message: false

library(data.table)
library(magrittr)
library(ggplot2)
library(RColorBrewer)
library(RplotterPkg)
library(RstocksPkg)
```

## The data

:::topic
History of posted dividends
:::

Get 12 years of historic data between the dates 2011-01-01 and 2022-12-31:

```{r}
#| warning: false
#| tbl-cap: Dividends posted for 9 NASDAQ companies from 2011 through 2022 (424 x 4)

symbols <- c("IBM","HPQ","TXN","CSCO","INTC","ORCL","AAPL","MSFT","QCOM")

dividends_dt <- RstocksPkg::get_dividends(
  symbols = symbols, 
  from = "2011-01-01", 
  to = "2022-12-31",
  warnings = FALSE
)


RplotterPkg::create_table(
  x = dividends_dt[1:6,],
  container_width_px = 500
)
```

Plot the historic data.

```{r}
#| fig-cap: Dividends Posted by 9 Attractive Companies from 2011 through 2022 

RplotterPkg::create_scatter_plot(
  df = dividends_dt,
  aes_x = "Date",
  aes_y = "Dividend",
  aes_fill = "Symbol",
  x_title = "Year",
  y_title = "Dividends($)",
  x_limits = as.Date(c("2011-01-01", "2022-12-31")),
  x_major_breaks = seq(as.Date("2011-01-01"), as.Date("2022-12-31"), by="1 year"),
  x_date_labels = "%Y",
  rot_y_tic_label = TRUE,
  y_limits = c(0.0, 2.0),
  y_major_breaks = seq(0.0,2.0,0.2),
  pts_size = 2.0,
  pts_stroke = 0.5,
  show_minor_grids = FALSE
)+
ggplot2::scale_fill_discrete(
  type = RColorBrewer::brewer.pal(n = 9, name = "Set1")
)
```

## Most recent dividend

1.  Get the most recent dividend posted:

```{r}
#| tbl-cap: Most recent dividend postings (9 x 4)

most_recent_df <- dividends_dt[Dividend > 0, .SD[.N], by = Symbol]

RplotterPkg::create_table(
  x = most_recent_df,
  container_width_px = 400
)
```

Plot the latest postings.

```{r, message=FALSE, warning=FALSE}
#| fig-cap: Most recent dividend postings occurring in 2022

RplotterPkg::create_scatter_plot(
  df = most_recent_df,
  aes_x = "Date",
  aes_y = "Dividend",
  aes_label = "Symbol",
  aes_label_size = 3,
  x_title = "Week (2022)",
  y_title = "Dividends($)",
  x_limits = as.Date(c("2022-10-01", "2022-12-31")),
  x_major_breaks = seq(as.Date("2022-10-01"), as.Date("2022-12-31"), by="1 week"),
  x_date_labels = "%b%e",
  rot_y_tic_label = TRUE,
  y_limits = c(0.0, 2.0),
  y_major_breaks = seq(0.0,2.0,0.2),
  pts_size = 1.2,
  show_minor_grids = FALSE,
  pts_color = "blue",
  pts_fill = "blue"
) 
```

## Total Annual Yield for 2022

The Total Annual Yield (TAY) is defined as the sum of total dividends in 2022 divided by the closing price on the first dividend date.

Find the first dividend for each company in 2022:

```{r}
first_dividend_2022_dt <- dividends_dt[Date > as.Date("2021-12-31"), .SD[1L], by = Symbol][order(-Dividend)] %>% 
data.table::setnames(old = "Dividend", new = "FirstDiv")
```

Find the sum of the total dividends in 2022:
```{r}
sum_dividend_2022_dt <- dividends_dt[Date > as.Date("2021-12-31"), lapply(.SD, sum), by = Symbol, .SDcols = "Dividend"][order(-Dividend)] %>% 
.[, .(Symbol, TotalDiv = Dividend)]  
```

Outer join both the "first" and "sum" data.tables.
```{r}
data.table::setkey(first_dividend_2022_dt, "Symbol")
data.table::setkey(sum_dividend_2022_dt, "Symbol")
tay_dividend_2022_dt <- first_dividend_2022_dt[sum_dividend_2022_dt]
```

Get the daily closing prices for the 9 companies.
```{r}
prices_dt <- RstocksPkg::get_stock_prices(
  symbols = tay_dividend_2022_dt$Symbol,
  periodicity = "daily",
  from = "2022-01-01",
  to = "2022-03-30"
) %>% 
  .[, .(Symbol, Date, Close)]
```

Outer join the closing prices with the `tay_dividends_2022_dt`.
```{r}
data.table::setkeyv(tay_dividend_2022_dt, cols = c("Symbol","Date"))
data.table::setkeyv(prices_dt, cols = c("Symbol","Date"))
tay_dividend_prices_dt <- prices_dt[tay_dividend_2022_dt]
```

Compute `TAY = (TotalDiv/Close) * 100`.
```{r}
tay_dividend_prices_dt <- tay_dividend_prices_dt[, TAY := round((TotalDiv/Close)*100,digits = 2)] %>% 
.[order(-TAY)]
```

Plot the TAY for the companies as a bar chart.

```{r}
#| fig-cap: Total Annual Yield for 2022

RplotterPkg::create_bar_plot(
  df = tay_dividend_prices_dt,
  aes_x = "Symbol",
  aes_y = "TAY",
  x_title = "Symbol",
  y_title = "Yield (%)",
  do_coord_flip = T,
  order_bars = "asc",
  rot_y_tic_label = T,
  bar_fill = "green",
  bar_alpha = 0.5,
  bar_labels = TRUE
)
```
